---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dev = "ragg_png",
  dpi = 300
)
```

# marquee

<!-- badges: start -->
[![R-CMD-check](https://github.com/thomasp85/marquee/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/thomasp85/marquee/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/thomasp85/marquee/branch/main/graph/badge.svg)](https://app.codecov.io/gh/thomasp85/marquee?branch=main)
<!-- badges: end -->

marquee is a markdown parser and renderer for the R graphics engine. It can be 
used to render rich text formatted as markdown (CommonMark) inside R graphics
such as ggplot2 or other graphics build on grid

## Installation

marquee is not yet on CRAN. In the meantime you can install the development 
version like so:

``` r
pak::pak("thomasp85/marquee")
```

## Examples
The main function of the package is `marquee_grob()` which creates a grob based
on markdown text and a style that can be rendered with grid:

```{r, fig.asp=2}
# Let's render this readme
readme <- paste(readLines("README.Rmd")[-seq_len(17)], collapse = "\n")

library(marquee)
library(grid)

fancy_style <- classic_style(
  body_font = "baskerville", 
  header_font = "marker felt",
  code_font = "fira code"
) |> 
  modify_style("cb", background = linearGradient(
    colours = c("lightblue", "white"),
    x1 = 0, y1 = 1, x2 = 0, y2 = 0
  ))
grob <- marquee_grob(readme, style = fancy_style)
grid.draw(grob)
```

(*The above is an image - go ahead and check*)

## Prior art
I would be remiss to not mention [gridtext](https://github.com/wilkelab/gridtext) 
and [ggtext](https://github.com/wilkelab/ggtext), both by Claus Wilke. These
packages aims to solve much the same as marquee, but work in a different way and
doesn't have the same powerful textshaping backend as marquee. Most notably from
a user perspective is perhaps that gridtext understands HTML to some degree, 
whereas marquee is oblivious to both HTML and CSS. Instead it supports the full
CommonMark spec with the plan to add support for custom span elements as well.
